Human Factors Influencing Visual Statistical Inference

### 2013-07-18 sociology_paper.tex
tweaking little things

\begin{table*}[hbtp] 
\centering 
\caption{Visual test statistics used in 10 different simulation experiments. The observers are asked different questions to answer while evaluating a lineup. All the experiment data are used to estimate the effect of demographic factors (DF) on visual inference while some are used to estimate the learning trend (LT) and location effect (LE) of actual data in the lineup as shown in the right most column.} 
\begin{tabular}{m{.5cm}m{2.6cm}m{2cm}m{5.5cm}m{3cm}} 
\hline\hline 
Serial & Experiment &  Test Statistic  & Lineup question & Designed to study\\ [0.5ex] % inserts table %heading 
\hline 
1  & Box plot & \begin{minipage}[t]{2cm} \begin{center}	\scalebox{0.12}{\includegraphics{stat_category1.pdf}} \end{center} \end{minipage} & Which set of box plots shows biggest vertical difference 
between group A and B? & DF\\
2 &  Scatter plot & \begin{minipage}[t]{2cm}  \begin{center} \scalebox{0.3}{\includegraphics{stat_beta_k1.png}} \end{center} \end{minipage} & Of the scatter plots below which one shows data that has steepest slope? & DF\\
  3 & Contaminated plot &\begin{minipage}[t]{2cm} \begin{center} \scalebox{0.5}{\includegraphics{stat_contaminated.pdf}} \end{center} \end{minipage} & Of the scatter plots below which one shows data that has steepest slope? & DF\\
 4 & Polar vs Cartesian & \begin{minipage}[t]{2cm} \begin{center}  \scalebox{0.32}{\includegraphics{stat_polar.pdf}} \end{center} \end{minipage} &Which plot is different?& DF\\
  5 & Hist vs density & \begin{minipage}[t]{2cm} \begin{center}  \scalebox{0.38}{\includegraphics{stat_density.pdf}} \end{center} \end{minipage} &In which plot is the blue group furthest to the right?&  DF and LT \\  
  6 & Violin vs boxplot & \begin{minipage}[t]{2cm} \begin{center}  \scalebox{0.35}{\includegraphics{stat_violin.pdf}} \end{center} \end{minipage}&  In which plot does the blue group look the most different from  the red group? &  DF and LT \\
  7 & Group separation & \begin{minipage}[t]{2cm} \begin{center}  \scalebox{0.4}{\includegraphics{stat_separation.pdf}} \end{center} \end{minipage} &Which of these plots has the most separation between the coloured groups? & DF and LT \\ 
  8 & Sine Illusion & \begin{minipage}[t]{2cm} \begin{center}  \scalebox{0.28}{\includegraphics{stat_sine.pdf}} \end{center} \end{minipage} & In what picture is the size of the curve most consistent? & DF\\ 
  9 & Gene expression &\begin{minipage}[t]{2cm} \begin{center}  \scalebox{0.45}{\includegraphics{stat_gene.pdf}} \end{center} \end{minipage} & In which of these plots is the green line the steepest, and the spread of the green points relatively small? & DF and LE\\ 
  10 & Test normality & \begin{minipage}[t]{2cm} \begin{center}  \scalebox{0.35}{\includegraphics{stat_normality.pdf}} \end{center} \end{minipage} &Which of these plots is most different from the others? & DF\\

@@

Based on human evaluation on a lineup a decision is made in visual statistical inference. While the visual test statistic is defined to have a direct influence on the observer to pick a plot based on the signal in the data, there are other human factors that may affect the observer performance. It is important to study the effect of those factors and examine the extent of their influence. This section presents a brief description of the factors that may affect the power of visual inference.

>>>>

\begin{table*}[hbtp] 
\centering 
\caption{Overview of 10 different Turk experiments, from where data was taken to study human factor effects. All of the experimental data was used to estimate the effect of demographic factors (DF) on visual inference while three were suitable for assessing learning trend (LT) and location effect (LE) was possible to assess using just one specially designed study.} 
\begin{tabular}{m{.5cm}m{2.6cm}m{2cm}m{5.5cm}m{3cm}} 
\hline\hline 
 & Experiment &  Test Statistic  & Lineup question & Used in study of\\ [0.5ex] % inserts table %heading 
\hline 
1  & Box plot & \begin{minipage}[t]{2cm} \begin{center}	\scalebox{0.12}{\includegraphics{stat_category1.pdf}} \end{center} \end{minipage} & Which set of box plots shows biggest vertical difference 
between group A and B? & DF\\
2 &  Scatter plot & \begin{minipage}[t]{2cm}  \begin{center} \scalebox{0.3}{\includegraphics{stat_beta_k1.png}} \end{center} \end{minipage} & Of the scatter plots below which one shows data that has steepest slope? & DF\\
  3 & Contaminated plot &\begin{minipage}[t]{2cm} \begin{center} \scalebox{0.5}{\includegraphics{stat_contaminated.pdf}} \end{center} \end{minipage} & Of the scatter plots below which one shows data that has steepest slope? & DF\\
 4 & Polar vs Cartesian & \begin{minipage}[t]{2cm} \begin{center}  \scalebox{0.32}{\includegraphics{stat_polar.pdf}} \end{center} \end{minipage} &Which plot is different?& DF\\
  5 & Hist vs density & \begin{minipage}[t]{2cm} \begin{center}  \scalebox{0.38}{\includegraphics{stat_density.pdf}} \end{center} \end{minipage} &In which plot is the blue group furthest to the right?&  DF and LT \\  
  6 & Violin vs boxplot & \begin{minipage}[t]{2cm} \begin{center}  \scalebox{0.35}{\includegraphics{stat_violin.pdf}} \end{center} \end{minipage}&  In which plot does the blue group look the most different from  the red group? &  DF and LT \\
  7 & Group separation & \begin{minipage}[t]{2cm} \begin{center}  \scalebox{0.4}{\includegraphics{stat_separation.pdf}} \end{center} \end{minipage} &Which of these plots has the most separation between the coloured groups? & DF and LT \\ 
  8 & Sine Illusion & \begin{minipage}[t]{2cm} \begin{center}  \scalebox{0.28}{\includegraphics{stat_sine.pdf}} \end{center} \end{minipage} & In what picture is the size of the curve most consistent? & DF\\ 
  9 & Gene expression &\begin{minipage}[t]{2cm} \begin{center}  \scalebox{0.45}{\includegraphics{stat_gene.pdf}} \end{center} \end{minipage} & In which of these plots is the green line the steepest, and the spread of the green points relatively small? & DF and LE\\ 
  10 & Test normality & \begin{minipage}[t]{2cm} \begin{center}  \scalebox{0.35}{\includegraphics{stat_normality.pdf}} \end{center} \end{minipage} &Which of these plots is most different from the others? & DF\\

@@

Because visual statistical inference relies on human evaluation, it is important to understand the factors that might affect results.  This section provides a brief description of the factors that may influence results.

%ased on human evaluation on a lineup a decision is made in visual statistical inference. While the visual test statistic is defined to have a direct influence on the observer to pick a plot based on the signal in the data, there are other human factors that may affect the observer performance. It is important to study the effect of those factors and examine the extent of their influence. This section presents a brief description of the factors that may affect the power of visual inference.
